---
title: "Distance-Decay Boxplot"
author: "Matthew S. Schechter"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This notebook describes the workflow for the distance-decay analysis of the TARA ocean prokaryotic, surface samples. We want to exploration the relationship between genetic distance and physical distance. Are samples more genetically similar if they are closer to eachother?

To reproduce this workflow upload `clstrs_tara_f1.RData` from figshare (10.6084/m9.figshare.5979658).

## Set environment
```{r}
# Load Libraries
library(tidyverse)
library(phyloseq)
library(data.table)
library(vegan)
library(RSQLite)
library(ggpubr)
library(GGally)
library(simba)
library(geosphere)
library(rworldmap)
library(scales)
library(sqldf)
library(RSQLite)
```

## Load data

### Component abundance data
```{r}
load("~/Desktop/MarMic/msc_thesis/data/clstrs_tara_f1.RData")
clstrs <- clstrs_tps_stats_f1
```

### Contextual data
```{r}
# Connect to SQLite databases to read in TARA contextual data.

# USAGE for RSQLite::dbConnect()
# - drv = typle of databse connection
# - dbname = path to database
db <- RSQLite::dbConnect(drv = SQLite(), dbname = "../../../databases/contextual_data.db")

# list contents of database
dbListTables(db)

# extract TARA Oceans contextual data from database
TARA_contex <- as_tibble(dbReadTable(db, "TARA_contex"))

# Disconnect from db
dbDisconnect(db)

# Filter contextual data for correct samples
contex <- TARA_contex %>% filter(sample_ID %in% clstrs$sample_ID)

# Add meren sites and temperature range classification
meren <- read_tsv("~/Desktop/MarMic/msc_thesis/meren/tom_tara_samples.tsv", col_names = TRUE) %>% rename(sample_ID = tara_sample) %>% separate(col = tom_sample, into = c("meren_region","meren_stuff"), sep = "_", remove = FALSE) %>% 
  dplyr::select(sample_ID, meren_region)

save(meren, file = "~/Desktop/MarMic/msc_thesis/data/meren.RData")

contex <- contex %>%
  mutate(t_range = ifelse(temperature < 19, "low",
                          ifelse(temperature >= 19 & temperature < 26,
                                 "medium", "high"))) %>%
  mutate(depth_category = case_when( grepl("SRF", sample_ID)  ~ 'SRF',
                                     grepl("DCM", sample_ID) ~ 'DCM',
                                     grepl("MES", sample_ID) ~ 'MES')) %>%
  left_join(meren)
```


## Prep data for analysis

Regions used by [Delmont et. al., 2017](https://www.biorxiv.org/content/early/2017/04/23/129791) were used to separate the TARA samples into different regions.

```{r}
# load abundance data to repeat analysis
# location on server "/bioinf/projects/megx/UNKNOWNS/Matt/TARA_clstrs_all_prok_meanprop.RData"
#load("~/Desktop/MarMic/msc_thesis/data/clstrs_tara_f1.RData")
#load("~/Downloads/TARA_contex (1).RData")

# Spread data into abundance matrix
clstrs <- clstrs_tps_stats_f1 %>%
  dplyr::select(sample_ID, component, abun) %>%
  spread(sample_ID, abun, fill = 0)

# make sample_ID rownames
clstrs <- as.data.frame(clstrs)
row.names(clstrs) <- clstrs$component
clstrs$component <- NULL

# Select component with mean proportion > 1e-5
comps_sel_meanprop <- clstrs_tps_stats_f1 %>%
  filter(mean_proportion >= 1e-5)

# Prep for phyloseq
clstrs <- as.matrix(clstrs)
TARA_contex_df <- contex
rownames(TARA_contex_df) <- TARA_contex_df$sample_ID

# Make phyloseq object
clstrs_physeq <- phyloseq(otu_table(clstrs, taxa_are_rows = TRUE), sample_data(TARA_contex_df))

#####################################################
# Cumulative Sum Scaling normalisation (CSS) function
#####################################################
# Make a cssTrans function
cssTrans <- function(f.physeq.p = f.physeq.p, norm = norm, log = log){
  require(metagenomeSeq)
  if (taxa_are_rows(f.physeq.p)) {
    f.physeq.p <- (f.physeq.p)
  }else{
    f.physeq.p <- t(f.physeq.p)
  }
  
  OTU <- as((otu_table(f.physeq.p, taxa_are_rows = TRUE)), "matrix")
  MGS <- newMRexperiment(
    counts = (OTU)
  )
  MGS <- cumNorm(MGS, p = cumNormStat(MGS))
  f.norm.p <- f.physeq.p
  otu_table(f.norm.p) <- otu_table((as.matrix(MRcounts(
    MGS,
    norm = norm,
    log = log,
    sl = median(unlist(normFactors(MGS)))
  ))), taxa_are_rows = T)
  return(f.norm.p)
}

# Let's CSS transform the data
clstrs_physeq_css <- cssTrans(clstrs_physeq, norm = TRUE, log = TRUE)

# Select high abundance clusters
clstrs_physeq_css_meanprop <- prune_taxa(comps_sel_meanprop$component, clstrs_physeq_css)

clstrs <- comps_sel_meanprop

# Filter contextual data for correct samples
contex <- TARA_contex %>% tbl_df() %>% dplyr::filter(sample_ID %in% clstrs$sample_ID)

# Add meren sites and temperature range classification
meren <- read_tsv("~/Desktop/MarMic/msc_thesis/meren/tom_tara_samples.tsv", col_names = TRUE) %>% rename(sample_ID = tara_sample) %>% separate(col = tom_sample, into = c("meren_region","meren_stuff"), sep = "_", remove = FALSE)

# Add the regions to contextual data
contex <- contex %>% left_join(meren)

# Separate the contextual data into the regions
# Atlantic
contex_atl <- contex %>% filter(meren_region == "ANW" | meren_region == "ANE" | meren_region == "ASW" | meren_region == "ASE") %>% mutate(meren_large_region = "atl") %>%
  filter(latitude > 0)

# Pacific
contex_pac <- contex %>% filter(meren_region == "PON" | meren_region == "POSW" | meren_region == "PSE") %>% mutate(meren_large_region = "pca")

# Indian
contex_in <- contex %>% filter(meren_region == "ION" | meren_region == "IOS") %>% mutate(meren_large_region = "in")

# Make 1 coordinate object
all_meren_coor <- bind_rows(contex_atl, contex_pac, contex_in) %>% 
  dplyr::select(sample_ID, latitude, longitude, meren_large_region)
  
# number of samples
n_samples <- clstrs  %>%
  dplyr::select(sample_ID) %>%
  unique() %>%
  count()

# If a component is found in all samples then it is ubiquitous
ubiq <- clstrs %>% group_by(component) %>% count %>% filter(n >= n_samples)

# If a component is NOT found in all samples then its is non-ubiquitous
nonubiq <- clstrs %>% group_by(component) %>% count %>% filter(n < n_samples)
```

## Distance decay calculations
```{r}
TARA_coords <- TARA_contex %>% 
  filter(sample_ID %in% clstrs$sample_ID) %>%
  dplyr::select(sample_ID, longitude, latitude) %>%
  arrange(desc(sample_ID)) %>% 
  as.data.frame() %>%
  column_to_rownames(var = "sample_ID")

TARA_dm_hav <- geosphere::distm(TARA_coords, fun=geosphere::distHaversine)
rownames(TARA_dm_hav) <- rownames(TARA_coords)
colnames(TARA_dm_hav) <- rownames(TARA_coords)
TARA_dm_hav_df <- broom::tidy(as.dist(TARA_dm_hav))

# All
clstrs_physeq_css_meanprop_bc_all <- vegan::vegdist(phyloseq:::veganifyOTU(clstrs_physeq_css_meanprop))
clstrs_physeq_css_meanprop_bc_all_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_all)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_all <- clstrs_physeq_css_meanprop_bc_all_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_all) <- c("loc1","loc2","value_dis","value_geo")

# EU
clstrs_comp_eu <- clstrs %>% filter(class == "eu") %>% .$component %>% unique
clstrs_physeq_css_meanprop_bc_eu <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_eu, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_eu_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_eu)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_eu <- clstrs_physeq_css_meanprop_bc_eu_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_eu) <- c("loc1","loc2","value_dis","value_geo")

## Ubiquitous
clstrs_comp_eu_ubi <- purrr::keep(clstrs_comp_eu, clstrs_comp_eu %in%  ubiq$component)
clstrs_physeq_css_meanprop_bc_eu_ubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_eu_ubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_eu_ubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_eu_ubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_eu_ubi <- clstrs_physeq_css_meanprop_bc_eu_ubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_eu_ubi) <- c("loc1","loc2","value_dis","value_geo")

## NON-Ubiquitous
clstrs_comp_eu_nonubi <- purrr::keep(clstrs_comp_eu, clstrs_comp_eu %in%  nonubiq$component)
clstrs_physeq_css_meanprop_bc_eu_nonubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_eu_nonubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_eu_nonubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_eu_nonubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_eu_nonubi <- clstrs_physeq_css_meanprop_bc_eu_nonubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_eu_nonubi) <- c("loc1","loc2","value_dis","value_geo")

# GU
clstrs_comp_gu <- clstrs %>% filter(class == "gu") %>% .$component %>% unique
clstrs_physeq_css_meanprop_bc_gu <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_gu, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_gu_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_gu)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_gu <- clstrs_physeq_css_meanprop_bc_gu_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_gu) <- c("loc1","loc2","value_dis","value_geo")

## Ubiquitous
clstrs_comp_gu_ubi <- purrr::keep(clstrs_comp_gu, clstrs_comp_gu %in%  ubiq$component)
clstrs_physeq_css_meanprop_bc_gu_ubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_gu_ubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_gu_ubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_gu_ubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_gu_ubi <- clstrs_physeq_css_meanprop_bc_eu_ubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_gu_ubi) <- c("loc1","loc2","value_dis","value_geo")

## NON-Ubiquitous
clstrs_comp_gu_nonubi <- purrr::keep(clstrs_comp_gu, clstrs_comp_gu %in%  nonubiq$component)
clstrs_physeq_css_meanprop_bc_gu_nonubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_gu_nonubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_gu_nonubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_gu_nonubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_gu_nonubi <- clstrs_physeq_css_meanprop_bc_gu_nonubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_gu_nonubi) <- c("loc1","loc2","value_dis","value_geo")

# unk
clstrs_comp_unk <- clstrs %>% filter(class == "eu" | class == "gu") %>% .$component %>% unique
clstrs_physeq_css_meanprop_bc_unk <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_unk, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_unk_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_unk)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_unk <- clstrs_physeq_css_meanprop_bc_unk_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_unk) <- c("loc1","loc2","value_dis","value_geo")

## Ubiquitous
clstrs_comp_unk_ubi <- purrr::keep(clstrs_comp_unk, clstrs_comp_unk %in%  ubiq$component)
clstrs_physeq_css_meanprop_bc_unk_ubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_unk_ubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_unk_ubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_unk_ubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_unk_ubi <- clstrs_physeq_css_meanprop_bc_unk_ubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_unk_ubi) <- c("loc1","loc2","value_dis","value_geo")

## NON-Ubiquitous
clstrs_comp_unk_nonubi <- purrr::keep(clstrs_comp_unk, clstrs_comp_unk %in%  nonubiq$component)
clstrs_physeq_css_meanprop_bc_unk_nonubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_unk_nonubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_unk_nonubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_unk_nonubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_unk_nonubi <- clstrs_physeq_css_meanprop_bc_unk_nonubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_unk_nonubi) <- c("loc1","loc2","value_dis","value_geo")

# known
clstrs_comp_known <- clstrs %>% filter(class == "k" | class == "kwp") %>% .$component %>% unique
clstrs_physeq_css_meanprop_bc_known <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_known, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_known_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_known)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_known <- clstrs_physeq_css_meanprop_bc_known_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_known) <- c("loc1","loc2","value_dis","value_geo")

## Ubiquitous

clstrs_comp_known_ubi <- purrr::keep(clstrs_comp_known, clstrs_comp_known %in%  ubiq$component)
clstrs_physeq_css_meanprop_bc_known_ubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_known_ubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_known_ubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_known_ubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_known_ubi <- clstrs_physeq_css_meanprop_bc_known_ubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_known_ubi) <- c("loc1","loc2","value_dis","value_geo")

## NON-Ubiquitous
clstrs_comp_known_nonubi <- purrr::keep(clstrs_comp_known, clstrs_comp_known %in%  nonubiq$component)
clstrs_physeq_css_meanprop_bc_known_nonubi <- vegan::vegdist(phyloseq:::veganifyOTU(prune_taxa(clstrs_comp_known_nonubi, clstrs_physeq_css_meanprop)))
clstrs_physeq_css_meanprop_bc_known_nonubi_df <- broom::tidy(as.dist(as.matrix(clstrs_physeq_css_meanprop_bc_known_nonubi)[rownames(TARA_coords),rownames(TARA_coords)]))
dis_gm_known_nonubi <- clstrs_physeq_css_meanprop_bc_known_nonubi_df %>% left_join(TARA_dm_hav_df, by = c("item1", "item2"))
colnames(dis_gm_known_nonubi) <- c("loc1","loc2","value_dis","value_geo")
```

## Choose which categories you want to show
```{r}
##########
# CHOOSE ONE!!!!!!!!!
###########

# Ubiquitous vs non-ubiquitous
all_dist <- bind_rows(#dis_gm_eu %>% mutate(class = "Environmental unknown"),
                      dis_gm_eu_ubi %>% mutate(class = "EU ubiquitous"),
                      dis_gm_eu_nonubi %>% mutate(class = "EU non-ubiquitous"),
                      #dis_gm_gu %>% mutate(class = "Genomic unknown"),
                      #dis_gm_unk %>% mutate(class = "Unknown"),
                      dis_gm_unk_ubi %>% mutate(class = "Unknown ubiquitous"),
                      dis_gm_unk_nonubi %>% mutate(class = "Unknown non-ubiquitous"),
                      #dis_gm_known %>% mutate(class = "Known"),
                      dis_gm_known_ubi %>% mutate(class = "Known ubiquotous"), 
                      dis_gm_known_nonubi %>% mutate(class = "Known non-ubiquotous"))

# Combined categories
all_dist_comb <- bind_rows(#dis_gm_eu %>% mutate(class = "Environmental unknown"),
                      #dis_gm_eu_ubi %>% mutate(class = "EU ubiquitous"),
                      #dis_gm_eu_nonubi %>% mutate(class = "EU non-ubiquitous"),
                      #dis_gm_gu %>% mutate(class = "Genomic unknown"),
                      dis_gm_unk %>% mutate(class = "Unknown"),
                      #dis_gm_unk_ubi %>% mutate(class = "Unknown ubiquitous"),
                      #dis_gm_unk_nonubi %>% mutate(class = "Unknown non-ubiquitous"),
                      dis_gm_known %>% mutate(class = "Known"),
                      #dis_gm_known_ubi %>% mutate(class = "Known ubiquotous"), 
                      #dis_gm_known_nonubi %>% mutate(class = "Known non-ubiquotous"),
                      dis_gm_all %>% mutate(class = "All"))

# Filter for regions
atlantic <- all_dist %>%
  filter(loc1 %in% contex_atl$sample_ID, loc2 %in% contex_atl$sample_ID)
pacific <- all_dist %>%
  filter(loc1 %in% contex_pac$sample_ID, loc2 %in% contex_pac$sample_ID)
indian <- all_dist %>%
  filter(loc1 %in% contex_in$sample_ID, loc2 %in% contex_in$sample_ID)

atlantic_comb <- all_dist_comb %>%
  filter(loc1 %in% contex_atl$sample_ID, loc2 %in% contex_atl$sample_ID)
pacific_comb <- all_dist_comb %>%
  filter(loc1 %in% contex_pac$sample_ID, loc2 %in% contex_pac$sample_ID)
indian_comb <- all_dist_comb %>%
  filter(loc1 %in% contex_in$sample_ID, loc2 %in% contex_in$sample_ID)
```

## Visualize
```{r}
# Plot distance-decay

p1 <- ggplot(atlantic, aes(x = (value_geo/1000), y = value_dis, color = class)) +
  geom_point(alpha = 0.5) + 
  stat_smooth() +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
        legend.position = "bottom") +
  scale_x_continuous(labels = scales::comma) +
  ylim(0, 1) +
  ggtitle("Atlantic") +
  xlab("Geographical distance (km)") +
  ylab("Bray-Curtis dissimmilarity")

p2 <- ggplot(pacific, aes(x = (value_geo/1000), y = value_dis, color = class)) +
  geom_point(alpha = 0.5) + 
  stat_smooth() +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
        legend.position = "bottom") +
  scale_x_continuous(labels = scales::comma) +
  ylim(0, 1) +
  ggtitle("Pacific") +
  xlab("Geographical distance (km)") +
  ylab("Bray-Curtis dissimmilarity")


p3 <- ggplot(indian, aes(x = (value_geo/1000), y = value_dis, color = class)) +
  geom_point(alpha = 0.5) + 
  stat_smooth() +
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
        legend.position = "bottom") +
  scale_x_continuous(labels = scales::comma) +
  ylim(0, 1) +
  ggtitle("Indian") +
  xlab("Geographical distance (km)") +
  ylab("Bray-Curtis dissimmilarity")

# Box-plot of BC distance
p1a <- ggboxplot(data = atlantic, x = "class", y = "value_dis", color = "class", add = "jitter") +
  ylim(0, 1) +
  labs(title = "") +
  xlab(label = "Category") +
  ylab(label = "Bray-Curtis dissimilarity")  +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "")

p2a <- ggboxplot(data = pacific, x = "class", y = "value_dis", color = "class", add = "jitter") +
  ylim(0, 1) +
  labs(title = "") +
  xlab(label = "Category") +
  ylab(label = "Bray-Curtis dissimilarity") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "")

p3a <- ggboxplot(data = indian, x = "class", y = "value_dis", color = "class", add = "jitter") +
  ylim(0, 1) +
  labs(title = "") +
  xlab(label = "Category") +
  ylab(label = "Bray-Curtis dissimilarity") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "")

ggarrange(p1, p2, p3, p1a, p2a, p3a, labels = c("A)", "B)", "C)", "D)", "E)", "F)"), nrow = 2, ncol = 3) %>%
  annotate_figure(top = text_grob("TARA prokaryotic surface distance-decay", color = "red", face = "bold", size = 20, hjust = 0.5))
```


## Map of regions
```{r}
map.world <- map_data(map="world")

ggplot() +
  geom_map(data=map.world, map=map.world, aes(map_id=region, x=long, y=lat)) +
  geom_point(data = all_meren_coor, aes(x=longitude, y=latitude, color = meren_large_region), size = 5) +
  theme_bw() +
  scale_size_manual(values = c(10,1)) +
  scale_shape_manual(values = c(4,19)) +
  theme(legend.position = "bottom",
        axis.text =  element_text()) +
  ggtitle(label = "Distance-decay surface samples")
  xlab("") +
  ylab("") +
  scale_x_continuous(breaks = c(-100, 0, 100))
  
  ggplot() +
  geom_map(data=map.world, map=map.world, aes(map_id=region, x=long, y=lat)) +
  geom_point(data = TARA_coor, aes(x=longitude, y=latitude), size = 2, color = 'darkblue') +
  theme_bw() +
  scale_size_manual(values = c(10,1)) +
  scale_shape_manual(values = c(4,19)) +
  theme(legend.position = "bottom",
        axis.text =  element_text()) +
  ggtitle(label = "Distance-decay surface samples")
  xlab("") +
  ylab("") +
  scale_x_continuous(breaks = c(-100, 0, 100))
```

