%% Chapter Template
\cleardoublepage{}
\newpage
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
\label{Acknowledgements}

\pagestyle{empty}
\fancyhead[RE]{}
\fancyhead[LO]{}

To begin, I would like to thank my supervisor Antonio who laid the ground work for me to become a successful PhD student. He has instilled critical thinking and methodical philosophy into my workflow. Where ever I end up I will truly benefit from these last 6 months with him as my supervisor.

A HUGE thanks to Chiara for all the support and not getting tired of me asking questions and barging into her office. 

Next, I would like to thank my two reviewers Frank Oliver and Pier. Thank you for your time and effort for providing me with constructive criticism on this thesis. 

Another thank you to Frank Oliver for allowing me to join the MGG group. I truly enjoyed working with this group and thank you to the whole MGG group for the support!

A separate thank you to Pier who provided with amazing explanations and guidance on my multivariate statistics. 

Thank you Henny for being my MSc partner in crime in MGG.

Thank you to my office mate Marcus Klimmek for all the laughs and drastically improving my German vocabulary in the most important ways.

Thank you to Gerard for working next to me during the beginning. It was fun learning Tidyverse with you!

Thank you MarMic class of 2021 for being an inspirational group of young scientists. I enjoyed learning and growing with you guys and cannot wait see how your careers excel!

Thank you to my coffee amigos Cora, David, and Andrea for all the great conversations and laughs!

Thank you to Christiane and Karl-Heinz for all the MarMic support.

Last but not least, many thanks to my Mother, Father, and Sister who encouraged me all along the way from the other side of the world. 